---
title: "Exploratory Data Analysis Template"
author: "Muuzaani Nkhoma"
date: "July 5, 2019"
output: html_document
---


Source: https://www.kaggle.com/uciml/student-alcohol-consumption
        https://cran.r-project.org/web/packages/dlookr/dlookr.pdf
        https://cran.r-project.org/web/packages/dlookr/vignettes/EDA.html


### Load Packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dlookr)                   # Exploratory Data Analysis
library(DataExplorer)             # Exploratory Data Analysis
library(GGally)                   # Parallel Coordinate Plots
library(cluster)                  # For clustering algorithms
library(factoextra)               # For visualization
library(dendextend)               # For comparing two dendrograms
library(magrittr)                 # For piping
library(kknn)                     # For K-Nearest Neighbors Analysis
library(e1071)                    # For Naive Bayes Algorithm
library(class)
library(MASS)                     # For Discriminant Analysis
library(klaR)                     # For Discriminant Analysis
 
```


## Read and prepare the data

### Dataset 1

#### Import dataset 1
```{r}
print("Start importing dataset")

d1=read_csv("Data/student-mat.csv", col_types = cols(
  school = col_factor(),
  sex = col_factor(),
  age = col_double(),
  address = col_factor(),
  famsize = col_factor(),
  Pstatus = col_factor(),
  Medu = col_factor(),
  Fedu = col_factor(),
  Mjob = col_factor(),
  Fjob = col_factor(),
  reason = col_factor(),
  guardian = col_factor(),
  traveltime = col_factor(),
  studytime = col_factor(),
  failures = col_factor(),
  schoolsup = col_factor(),
  famsup = col_factor(),
  paid = col_factor(),
  activities = col_factor(),
  nursery = col_factor(),
  higher = col_factor(),
  internet = col_factor(),
  romantic = col_factor(),
  famrel = col_factor(),
  freetime = col_factor(),
  goout = col_factor(),
  Dalc = col_factor(),
  Walc = col_factor(),
  health = col_factor(),
  absences = col_double(),
  G1 = col_double(),
  G2 = col_double(),
  G3 = col_double()
))

dim(d1)
ncol(d1)
nrow(d1)
names(d1)

print("Finished importing dataset")

```

##View first observations

```{r}
head(d1)

```

#### View last observations
```{r}
tail(d1)

```



```{r}

str(d1)

```



```{r}

d1_list <- list(d1)
plot_str(d1_list)

```


##### Descriptive Statistics


```{r}

introduce(d1)

```



```{r}

plot_intro(d1)

```


```{r}

summary(d1)

```





### CONTINUOUS VARIABLES   


#### UNIVARIATE ANALYSIS


```{r}

find_class(d1, type = "numerical",  index = FALSE)

```



```{r}

d1_num <- dplyr::select(d1, age, absences, G1, G2, G3)
  
```



##### Graphical Analysis



```{r}

plot_histogram(d1)

```



```{r}

plot_density(d1)

```



```{r}

plot_qq(d1)

```


```{r}

plot_normality(d1)

```




```{r}

ggplot(d1, aes(x = factor(0), y = absences)) + 
  geom_boxplot() + xlab("") + labs(title = "                                    Boxplot of First Period Grade")
  scale_x_discrete(breaks = NULL)

```




```{r}

boxplot(d1$G1, main = "Boxplot of First Period Grade", xlab = "Grade", ylab = "Count")

```




##### Non-graphical Analysis

```{r}

describe(d1)

```



```{r}

diagnose_numeric(d1)

```







#### BIVARIATE ANALYSIS



##### Graphical Analysis

```{r}

plot_correlate(d1)

```



```{r}

pairs(d1_num)

```



```{r}

plot_correlation(d1, type = "c")

```



```{r}

plot_scatterplot(d1, by = "G3")

```



```{r}

plot_boxplot(d1, by = "G3")

```



##### Non-graphical Analysis



```{r}

cor(d1_num)

```






### CATEGORICAL VARIABLES

#### UNIVARIATE ANALYSIS

```{r}

find_class(d1, type = "categorical2",  index = FALSE)

```




##### Non-graphical Analysis

```{r}

diagnose_category(d1)

```






#### BIVARIATE ANALYSIS



##### Graphical Analysis

```{r}

plot_bar(d1)

```



##### Non-graphical Analysis

Contingency Tables

```{r}

d1.table <- table(d1$failures)
d1.table

```



##### Graphical Analysis

Mosaic Plot

```{r}

mosaicplot(d1.table, main = deparse(substitute(d1.table)))

```





#### Missing Values

```{r}

plot_missing(d1)

```



#### OUTLIERS

```{r}

find_outliers(d1)
plot_outlier(d1)

```




### MULTIVARIATE ANALYSIS


#### PARALLEL COORDINATE PLOTS

```{r}

ggparcoord(d1, columns = c(3, 30:33), groupColumn = "Medu")

```



### PRINCIPAL COMPONENT ANALYSIS

```{r}

plot_prcomp(d1_num)

```


#### Biplot

```{r}

biplot(prcomp(d1_num))

```



#### Screeplot

```{r}

screeplot(prcomp(d1_num), type = "lines")

```



## CLUSTER ANALYSIS

### CLUSTERING METHODS

#### HIERARCHICAL CLUSTERING  
Note that agglomerative clustering is good at identifying small clusters. Divisive hierarchical clustering is good at identifying large clusters.


#### Distance Matrix

```{r}
# Dissimilarity matrix
distanceMatrix <- dist(d1_num, method = "euclidean")

fviz_dist(distanceMatrix, 
   gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07"))
```


#### Agglomerative Hierarchical Clustering

```{r}

# Hierarchical clustering using Complete Linkage
hc1 <- hclust(distanceMatrix, method = "complete" )

# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)

```


##### Including Agglomerative Coefficient

```{r}

# Compute with agnes
hc2 <- agnes(d1_num, method = "complete")

# Agglomerative coefficient
hc2$ac

# Plot the obtained dendrogram
pltree(hc2, cex = 0.6, hang = -1)

```
##### Model Assessment

```{r}
# methods to assess
m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")

# function to compute coefficient
ac <- function(x) {
  agnes(d1_num, method = x)$ac
}

map_dbl(m, ac)

```


##### Divisive Hierarchical Clustering


```{r}
# compute divisive hierarchical clustering
hc4 <- diana(d1_num)

# Divise coefficient; amount of clustering structure found
hc4$dc

# plot dendrogram
pltree(hc4, cex = 0.6, hang = -1, main = "Dendrogram of diana")
```

#### PARTITIONING CLUSTERING

##### Determining optimal number of clustering

##### Gap Statistic

```{r}

fviz_nbclust(d1_num, kmeans, method = "gap_stat")
```



##### Elbow Method

```{r}
fviz_nbclust(d1_num, FUN = hcut, method = "wss")
```


```{r}

set.seed(123)

d1.km <- kmeans(d1_num, 3, nstart = 25)

```


##### Plot Clusters

```{r}
# Visualize
library("factoextra")
fviz_cluster(d1.km, data = d1_num,
             ellipse.type = "convex",
             palette = "jco",
             ggtheme = theme_minimal())

```


```{r}

fviz_cluster(d1.km, data = d1_num)

```





```{r}

d1.km

```




```{r}

d1_Cluster <- cbind(d1, d1.km$cluster) %>% # combine the original data and the clusters
  mutate(Cluster = as_factor(d1.km$cluster)) %>%
  dplyr::select(-"d1.km$cluster")

  
```



```{r}

ggparcoord(d1_Cluster, columns = c(3, 30:33), groupColumn = "Cluster", scale = "std")


```





## PREDICTIONS

#### TARGET VARIABLE CATEGORICAL

Set target variable as categorical

```{r}

categ <- target_by(d1, school)

```



#### NUMERIC PREDICTORS and CATEGORICAL TARGET

```{r}

cat_num <- relate(categ, G3)
cat_num

```


```{r}

cat_num1 <- relate(categ, c(G1, G2, G3))
cat_num1

```


```{r}

summary(cat_num)

```


```{r}

plot(cat_num)

```




#### CATEGORICAL PREDICTORS and CATEGORICAL TARGET

```{r}

cat_cat <- relate(categ, Medu)
cat_cat

```



```{r}

plot(cat_cat)

```




#### TARGET VARIABLE NUMERICAL

Set target variable as categorical

```{r}

num <- target_by(d1, G3)

```



#### NUMERIC PREDICTORS and NUMERICAL TARGET

```{r}

num_num <- relate(num, age)
num_num

```


```{r}

summary(num_num)

```



```{r}

plot(num_num)

```




#### CATEGORICAL PREDICTORS and NUMERICAL TARGET

```{r}

num_cat <- relate(num, Medu)
num_cat

```



```{r}

summary(num_cat)

```



```{r}

plot(num_cat)

```






## CREATING AND SAVING AN R OBJECT INTO A CSV FILE

```{r}

write.table(d1_Cluster, "Data/EDA Example.csv", sep = ",", row.names = F) # save as a .csv file

```






